Synthesis of Rules for Network Intrusion Detectors

### 2019-10-25 paper/paper.tex
another pass

\subsection{Step 2: Rule Minimization}

@@

@@

%% \renewcommand{\labelitemi}{}
%% \setitemize[0]{leftmargin=20pt,itemindent=-20pt}
%% \begin{itemize}
%% \end{itemize}

%% \renewcommand{\labelitemi}{}
%% \setitemize[0]{leftmargin=0pt,itemindent=0pt}
%% \begin{itemize}
%%   \item{\textbf{RQ1.}~How precise is \tname?}
%%   \item{\textbf{RQ2.}~How efficient is \tname?}
%% \end{itemize}

This section reports precision and efficiency of \tname.

@@

We evaluated \tname{} on a set of attacks covering each of the four
protocols supported by Suricata and attacks with different
characteristics. Table~\ref{table:attacks} shows the attacks we
analyzed and the rule of Suricata that captures the attack. Column
``Name'' shows the name of the attack, column ``\#'' indicates whether
single or multiple messages are required to manifest the attack,
column ``Protocol'' shows the protocol name, column ``Description''
summarizes the effect of the attack, if successfull, and column
``Golden Rule'' shows the relevant parts in the options section of the
Suricata rule that captures the attack.

\begin{table*}[t!]
  \caption{\label{table:attacks}Characterization of attacks analyzed
    and Golden Rule of Suricata (reference) to capture it.\Mar{please,
  show me the Golden Rule, \ie{} the part that is relevant of each rule (for
  capturing the negative traffic precisely)}}  
  \centering
  \begin{tabular}{lllll}
    \toprule
    \multicolumn{1}{c}{Name} & \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Protocol} & \multicolumn{1}{c}{Description} & \multicolumn{1}{c}{Golden Rule} \\
    \midrule     
    Ping Scan & single & ICMP & Discovers active hosts in a network &  dsize:0; itype:8; \\
    Black Nurse & multiple & ICMP & Denial of Service & itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;\\
    Ping Flood  & multiple & ICMP & Denial of Service & itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;\\  
    Port Scan TCP & single & TCP & Discovers open ports in a host & flags: A; ack: 0;dsize: 0; \\
    SYN Flood & multiple & TCP & Denial of Service & flags: S,12; threshold: type both, track by\_dst, count 5000, seconds 5;\\
    UDP Flood & multiple & UDP & Denial of Service & fragbits:M; threshold: type both, track by\_dst, count 5000, seconds 5; \\
    \bottomrule
  \end{tabular}
\end{table*}

\subsection{Setup}

Note from Figure~\ref{fig:overview} that the negative traffic is an
input to the technique. The positive (non malicious) traffic, in
contrast, is not. \tname{} uses public databases storing different
kinds of positive traffic for all protocols covered by Suricata. We
used two datasets of positive traffic. The Bigflows.pcap dataset is
part of the TcpReplay~\cite{tcpreplay} open source project. This
dataset includes real network traffic on a busy private network access
point to the Internet. According to the TcpReplay web site ``This
capture is much larger and has a smaller average packet size than the
previous capture (smallFlows.pcap). It also has many more flows and
different applications.''. In addition to the Bigflows.pcap dataset,
we also used datasets of normal traffic made available by Stratosphere
Lab~\cite{stratosphere-normal}.

@@

\Luc{entrada: utilizamos a ferramenta hping3 para executar o ataque em um alvo arbitrario dentro da rede (n ha necessidade de haver um alvo real para executar o hping3) e capturamos o trafego gerado com o wireshark. o comando utilizado foi:\\*
\texttt{\# hping3 -d 80 -w 64 -S -p 80 --flood --rand-source 192.138.1.115} \\*
que envia pacotes de 80 bytes (-d 80), com a flag SYN habilitada (-S), tamanho de janela de 64 bytes (-w 64), direcionados a porta 80 (-p 80). A flag --flood indica q o hping3 vai enviar os pacotes o mais rápido possível. a flag --rand-source indica que cada pacote vai ter um ip de origem aleatorio. 192.138.1.115 é o alvo.\\*
O trafego armazenado contem 40 mil pacotes SYN, enviados no intervalo de 1 segundo,  pegamos uma amostra de 20 desses pacotes para representar o flood e servir como entrada.\\*
O wireshark armazena os pacotes no formato pcap, utilizamos a ferramenta tcpreplay para reproduzir o trafego com o comando:\\*
\texttt{\# tcpreplay --pps 200 --intf1=wlp2s0 Datasets/syn-flood-sample.pcap}\\*
Nesse caso, estamos enviando os pacotes armazenados em \texttt{Datasets/syn-flood-sample.pcap} a uma taxa de 200 pacotes por segundo (\texttt{-pps 200}) utilizando a interface wlp2s0 (\texttt{--intf1 wlp2s0})\\* 
Há diferenças entre os pacotes na opção ack.\\* 
Por que 20? Apenas para gerar a regra mais rapidamente, poderia ser qualquer valor maior ou menor. No final, o parâmetro \texttt{count} da opção \texttt{threshold} vai ser alterado pelo adm da rede.\\*
Intervalo de reprodução: 0.005 segundo

}

\Fix{any other detail we should consider?}

>>>>

\subsection{Step 2: Rule Minimization}
\label{sec:minimization}

@@

\vspace{1ex}
\noindent\textbf{Implementation details.} \Fix{discuss function isFit}

@@

This section evaluates the precision and efficiency of \tname{}.

@@

Table~\ref{table:attacks} shows the attacks we analyzed and the
options of the rule from the Emerging Threats Open
Ruleset~\cite{emerging-threats-open} that captures the attack. Column
``Name'' shows the name of the attack, column ``Protocol'' shows the
protocol name, and column ``Ground Truth'' shows the options of the
rule that captures the attack.

\begin{table*}[h]
  \caption{\label{table:attacks}List of attacks.\Mar{Guilherme, por
      favor, adicione mais ataques aqui...}}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{Protocol} &
    \multicolumn{1}{c}{Groud Truth} \\
    \midrule     
    Ping Scan \MyComment{& single} & ICMP \MyComment{& Discovers
      active hosts in a network} & \scriptsize{dsize:0; itype:8;} \\
    Black Nurse \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;}\\
    Ping Flood  \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;}\\  
    Port Scan \MyComment{& single} & TCP \MyComment{& Discovers open ports in a host} & \scriptsize{flags: A; ack: 0;dsize: 0;} \\
    SYN Flood \MyComment{& multiple} & TCP \MyComment{& Denial of Service} & \scriptsize{flags: S,12; threshold: type both, track by\_dst, count 5000, seconds 5;}\\
    UDP Flood \MyComment{& multiple} & UDP \MyComment{& Denial of Service} & \scriptsize{fragbits:M; threshold: type both, track by\_dst, count 5000, seconds 5;} \\
    \bottomrule
  \end{tabular}
\end{table*}

\subsection{Dataset of Benign Traffic}

Note from Figure~\ref{fig:overview} that the positive traffic is an
input to the technique. The negative non-malicious traffic, however,
is not. \tname{} uses public databases storing different kinds of
positive traffic for all protocols covered by Suricata. We used two
datasets of positive traffic. The Bigflows.pcap dataset is part of the
TcpReplay~\cite{tcpreplay} open source project. This dataset includes
real network traffic on a busy private network access point to the
Internet. According to the TcpReplay web site ``This capture is much
larger and has a smaller average packet size than the previous capture
(smallFlows.pcap). It also has many more flows and different
applications.''. In addition to the Bigflows.pcap dataset, we also
used datasets of normal traffic made available by Stratosphere
Lab~\cite{stratosphere-normal}.

\subsection{Methodology}

\tname{} expects (a) message(s) reproducing an attack on input. As
previously discussed, those messages could be directly provided by an
anomaly-based \nids\ tool. To evaluate \tname{}, however, we created
those messages ourselves. For all attacks listed on
Table~\ref{table:attacks}, we found reproduction scripts available on
the web. With the script to reproduce the attack, we proceeded as
follows: 1)~we instantiated the script on a given target machine,
2)~we executed the script, 3)~we monitored the traffic with the
wireshark network monitor~\cite{wireshark-net-monitor}, and 4) we
simulated the delivery of the corresponding network messages by
directly executing Suricata on the ``pcap'' files that wireshark
produced in the previous step. It is worth noting that Suricata is
invoked multiple times in the process of minimizing the rule (see
Section~\ref{sec:minimization}) and that processing pcap files is much
more efficient compared to sending and processing network messages. To
illustrate, the SYN Flood attack was reproduced with the hping3
tool~\cite{hping3} using the following command \CodeIn{hping3 -d 80 -w
  64 -S -p 80 --flood --rand-source <IP>}.

@@

%% \Luc{entrada: utilizamos a ferramenta hping3 para executar o ataque em um alvo arbitrario dentro da rede (n ha necessidade de haver um alvo real para executar o hping3) e capturamos o trafego gerado com o wireshark. o comando utilizado foi:\\*
%% \texttt{\# hping3 -d 80 -w 64 -S -p 80 --flood --rand-source 192.138.1.115} \\*
%% que envia pacotes de 80 bytes (-d 80), com a flag SYN habilitada (-S), tamanho de janela de 64 bytes (-w 64), direcionados a porta 80 (-p 80). A flag --flood indica q o hping3 vai enviar os pacotes o mais rápido possível. a flag --rand-source indica que cada pacote vai ter um ip de origem aleatorio. 192.138.1.115 é o alvo.\\*
%% O trafego armazenado contem 40 mil pacotes SYN, enviados no intervalo de 1 segundo,  pegamos uma amostra de 20 desses pacotes para representar o flood e servir como entrada.\\*
%% O wireshark armazena os pacotes no formato pcap, utilizamos a ferramenta tcpreplay para reproduzir o trafego com o comando:\\*
%% \texttt{\# tcpreplay --pps 200 --intf1=wlp2s0 Datasets/syn-flood-sample.pcap}\\*
%% Nesse caso, estamos enviando os pacotes armazenados em \texttt{Datasets/syn-flood-sample.pcap} a uma taxa de 200 pacotes por segundo (\texttt{-pps 200}) utilizando a interface wlp2s0 (\texttt{--intf1 wlp2s0})\\* 
%% Há diferenças entre os pacotes na opção ack.\\* 
%% Por que 20? Apenas para gerar a regra mais rapidamente, poderia ser qualquer valor maior ou menor. No final, o parâmetro \texttt{count} da opção \texttt{threshold} vai ser alterado pelo adm da rede.\\*
%% Intervalo de reprodução: 0.005 segundo

%% }

%% \Fix{any other detail we should consider?}
